Fit by design: Developing substrate-specific seed mixtures for functional dike grasslands

### 2023-02-28 manuscript.Rmd
Update manuscript.Rmd

MB is funded by a doctoral scholarship of the German Federal Environmental Foundation (DBU) (No. 20021/698). The establishment of the experiment and the vegetation surveys were financed by the WIGES GmbH in the years 2018--2020 (No. 80 002 312).

@@

Characteristics of the substrates used for the sowing experiment on river dikes. Soil samples of the three substrates were analyzed for the fraction \< 2 mm. Soil texture was classified according to the 'Bodenkundliche Kartieranleitung' [@BodenkundlKartieranl.2005]. The pH was measured in CaCl~2~ solution. Plant available phosphorus and potassium were measured in a calcium acetate-lactate extract and magnesium in CaCl~2~ extract. For calculating CaCO~3~, a sub-sample was annealed at 550 °C and the measured C amount multiplied with 8.33. To calculate total N and the C/N ratio, a sub-sample was incinerated at 1000 °C. Lt3 = medium clayey loam; Ls4 = strong sandy loam; Sl3 = medium loamy sand; Sl4 = strong loamy sand.

@@

Local setting and design of the multifactorial experiment on grassland sowing at dikes. The experiment was located on a dike at river Danube in SE Germany. The 288 plots were allocated in six blocks (white squares on upper photo) and on the north and south slope (central photo) [Both aerial photos: @bayerischevermessungsverwaltung2023]. Four treatments were conducted: sand admixture, substrate depth, seed density, and seed mixture types H and D (hay meadows, dry grasslands). The western half of a block was with a shallow substrate depth and within this half the substrates had different sand admixtures. The photo on the bottom shows the northern slope of one block in 2021, four years after sowing (Photo: Markus Bauer).

@@

```{r fig.id = "effects", fig.asp = 0.3030, fig.width = 6.49, fig.cap = "The posterior density distributions (grey) of the treatments calculated over all four surveyed years of sown grassland communities at a river dike. Furthermore, the median and the 66% and 95% credible intervals are shown: the persistence of sown species (A), the Favourable Conservation Status (FCS; B) and the successional distance to reference (recovery; C). The FCS is the ratio of target species and non-target species. Note that the zero lines indicate that both levels have equal values. This means for instance that hay meadows are closer to their reference than dry grasslands  (C)."}
knitr::include_graphics(
  here::here("outputs", "figures", "figure_2_800dpi_16.5x5cm.tiff")
  )
```

@@

```{r fig.id = "persistence", fig.asp = 0.4848, fig.width = 6.49, fig.cap = "Persistence (1-*B*~sor~) of sown species of sown grassland communities at a river dike. The plots had substrates with different sand admixtures and were sown with two different seed mixture types. Per plot, twenty species were sown. Shown are the medians and 95% credible intervals of the posterior distributions which were derived from a Bayesian linear mixed-effects model (BLMM)."}
knitr::include_graphics(
  here::here(
    "outputs", "figures", "figure_3_persistence_epred_800dpi_16.5x8cm.tiff"
    )
  )
```

@@

```{r fig.id = "fcs", fig.asp = 0.4848, fig.width = 6.49, fig.cap = "The Favourable Conservation Status (FCS) of sown grassland communities at a river dike. The plots had substrates with different sand admixtures and were sown with two different seed mixtures. The zero line indicate that target and non-target species are balanced. Positive values indicate that there are more target species. Shown are the medians and 95% credible intervals of the posterior distributions which were derived from a Bayesian linear mixed-effects model (BLMM)."}
knitr::include_graphics(
  here::here("outputs", "figures", "figure_4_fcs_epred_800dpi_16.5x8cm.tiff")
  )
```

@@

```{r fig.id = "recovery", fig.asp = 0.4848, fig.width = 6.49, fig.cap = "The recovery or successional distance (*d~jt,0~*) of sown grassland communities at a river dike in comparison to the respective reference sites (hay meadows or dry calcareous grasslands). The plots had substrates with different sand admixtures and were sown with two different seed mixtures. The zero lines indicate the mean position of the reference sites for each habitat type on the NMDS axis 1 (Figure 7). The grey area marks the standard deviation of the position of the reference sites. Shown are the medians and 95% credible intervals of the posterior distributions which were derived from a Bayesian linear mixed-effects model (BLMM)."}
knitr::include_graphics(
  here::here(
    "outputs", "figures", "figure_5_recovery_epred_800dpi_16.5x8cm.tiff"
    )
  )
```

@@

The species composition of sown experimental plots on a river dike over time and in comparison with reference sites and the seed mixtures. Both expositions and both seed mixture types are shown in separated panels. The NMDS was based on Sørensen dissimilarity and data of 288 plots which were observed over four years after sowing in 2018 (circles). These experimental plots were compared with the seed mixtures (black squares) and 98 positive and negative reference plots (filled symbols) from older dike grasslands in the surroundings [@bauer2023] and six plots from sPlotOpen [@Sabatini.2021]. The circles show the standard error of the groups. 2D-stress: 0.21.

>>>>

MB was funded by a doctoral scholarship of the German Federal Environmental Foundation (DBU) (No. 20021/698). The establishment of the experiment and the vegetation surveys were financed by the WIGES GmbH in the years 2018--2020 (No. 80 002 312).

@@

Characteristics of the substrates used for the sowing experiment on river dikes. Soil samples of the three substrates were analyzed for the fraction \< 2 mm. The soil texture was classified according to the 'Bodenkundliche Kartieranleitung' [@BodenkundlKartieranl.2005]. The pH was measured in CaCl~2~ solution. Plant available phosphorus and potassium were measured in a calcium acetate-lactate extract and magnesium in a CaCl~2~ extract. For calculating CaCO~3~, a sub-sample was annealed at 550 °C and the measured C amount multiplied with 8.33. To calculate total N and the C/N ratio, a sub-sample was incinerated at 1000 °C. Lt3 = medium clayey loam; Ls4 = strong sandy loam; Sl3 = medium loamy sand; Sl4 = strong loamy sand.

@@

Local setting and design of the multifactorial experiment on grassland sowing at dikes. The experiment was located on a dike at the Danube River in SE Germany. The 288 plots were allocated in six blocks (white squares on the upper photo) and on the north and south slope (central photo) [Both aerial photos: @bayerischevermessungsverwaltung2023]. Four treatments were conducted: sand admixture, substrate depth, seed density, and seed mixture types H and D (hay meadows, dry grasslands). The western half of a block was with a shallow substrate depth and within this, half of the substrates had different sand admixtures. The photo on the bottom shows the northern slope of one block in 2021, four years after sowing (Photo: Markus Bauer).

@@

```{r fig.id = "effects", fig.asp = 0.3030, fig.width = 6.49, fig.cap = "The posterior density distributions (grey) of the treatments calculated over all four surveyed years of sown grassland communities at a river dike. Furthermore, the median and the 66% and 95% credible intervals are shown: the persistence of sown species (A), the Favourable Conservation Status (FCS; B) and the successional distance to reference (recovery; C). The FCS is the ratio of target species and non-target species. Note that the zero lines indicate that both levels have equal values. This means, e.g., that hay meadows are closer to their reference than dry grasslands  (C)."}
knitr::include_graphics(
  here::here("outputs", "figures", "figure_2_800dpi_16.5x5cm.tiff")
  )
```

@@

```{r fig.id = "persistence", fig.asp = 0.4848, fig.width = 6.49, fig.cap = "Persistence (1-*B*~sor~) of sown species of sown grassland communities at a river dike. The plots had substrates with different sand admixtures and were sown with two different seed mixture types. Twenty species were sown per plot. Shown are the medians and 95% credible intervals of the posterior distributions, which were derived from a Bayesian linear mixed-effects model (BLMM)."}
knitr::include_graphics(
  here::here(
    "outputs", "figures", "figure_3_persistence_epred_800dpi_16.5x8cm.tiff"
    )
  )
```

@@

```{r fig.id = "fcs", fig.asp = 0.4848, fig.width = 6.49, fig.cap = "The Favourable Conservation Status (FCS) of sown grassland communities at a river dike. The plots had substrates with different sand admixtures and were sown with two different seed mixtures. The zero line indicates that target and non-target species are balanced. Positive values indicate that there are more target species. Shown are the medians and 95% credible intervals of the posterior distributions, which were derived from a Bayesian linear mixed-effects model (BLMM)."}
knitr::include_graphics(
  here::here("outputs", "figures", "figure_4_fcs_epred_800dpi_16.5x8cm.tiff")
  )
```

@@

```{r fig.id = "recovery", fig.asp = 0.4848, fig.width = 6.49, fig.cap = "The recovery or successional distance (*d~jt,0~*) of sown grassland communities at a river dike in comparison to the respective reference sites (hay meadows or dry calcareous grasslands). The plots had substrates with different sand admixtures and were sown with two different seed mixtures. The zero lines indicate the mean position of the reference sites for each habitat type on the NMDS axis 1 (Figure 7). The grey area marks the standard deviation of the position of the reference sites. Shown are the medians and 95% credible intervals of the posterior distributions, which were derived from a Bayesian linear mixed-effects model (BLMM)."}
knitr::include_graphics(
  here::here(
    "outputs", "figures", "figure_5_recovery_epred_800dpi_16.5x8cm.tiff"
    )
  )
```

@@

The species composition of sown experimental plots on a river dike over time and in comparison with reference sites and the seed mixtures. Both expositions and both seed mixture types are shown in separated panels. The NMDS was based on the Sørensen dissimilarity and data of 288 plots observed over four years after sowing in 2018 (circles). These experimental plots were compared with the seed mixtures (black squares) and 98 positive and negative reference plots (filled symbols) from older dike grasslands in the surroundings [@bauer2023] and six plots from sPlotOpen [@Sabatini.2021]. The circles show the standard error of the groups. 2D-stress: 0.21.
